Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-04 WSSSPE2_report/WSSSPE2_review.tex
working in intro, style, credit section

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
\newcommand{\gabnote}[1]{ {\textcolor{cyan}    { ***Gabrielle:     #1 }}}
\newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
\newcommand{\manishnote}[1]{  {\textcolor{violet}     { ***Manish: #1 }}}
\newcommand{\davidnote}[1]{  {\textcolor{darkgreen}      { ***David: #1 }}}
\newcommand{\colinnote}[1]{ {\textcolor{red}    {***Colin #1 }}}

@@

\begin{abstract} \todo{TBD due date above}       
This technical report discusses the submission and peer-review process used by
the Second Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE2) and the results of that process. It is intended to record
both the alternative submission and program organization model used by WSSSPE2
as well as the papers associated with the workshop that resulted from that
process.
\end{abstract}

@@

\todo{update this section with intro to full report - currently just intro for pre-workshop stuff. Link to live notes -
\url{http://tinyurl.com/q6ew45v}}
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Submissions, Peer-Review, and Peer-Grouping}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{shorten this section? - cite~\cite{WSSSPE2-pre-report} somewhere here}

The workshop call for papers included the following areas of interest:
\begin{quote}
\begin{itemize} 
\renewcommand{\labelenumi}{\textbf{\theenumi}.}
\setlength{\rightmargin}{1em}
\item defining software sustainability in the context of science and engineering
software
\begin{itemize}
\item how to evaluate software sustainability
\end{itemize}

@@

Based on the goal of encouraging a wide range of submissions from those involved
in software practice, ranging from initial thoughts and partial studies to
mature deployments, but focusing on papers that are intended to lead to changes,
the organizers wanted to make submission as easy as possible. The call for
papers stated:

@@

Finally, since some of the papers were not represented by cards in the process,
they were not placed in groups by the peer-grouping system, so the authors of
these papers were asked which group seemed the best for their papers, and those
papers were then placed in those groups. The later sections of this paper that
discuss the breakout groups include a list of the papers associated with each
group.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes} \label{keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Lead: Choi. Volunteers welcome---contact lead}

@@

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of Mozilla Science talked about her organization's
work and policy to enable and support sustainable and reproducible scientific
research. Our second keynote speaker was Neil Chue Hong of Software
Sustainability Institute. He shined the light on how scientific software is
prevalently driving advances in many sciences and engineering fields. Following
the keynotes, participants broke into groups to define and discuss the crucial
notion ``sustanability'' spawned by the speeches of our keynote speakers as well
as the theme of our workshop.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exploring Sustainability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Colin to lead writing of this section.}

\href{http://tinyurl.com/mpbhvyb}{Google doc notes}
%https://docs.google.com/a/uchicago.edu/document/d/10XkshP3YjXnA5JQgP-UFHbOxshWczzehWGSo6V7bNH4/edit

@@

@@

\href{http://tinyurl.com/pn4eq8z}{Google doc notes}
%https://docs.google.com/a/uchicago.edu/document/d/1_q0jmiEPNFRGAEtGLk_QxQ1vVokgTDYruTGURowNGkY/edit

@@

@@

\href{http://tinyurl.com/k8ruyn9}{Google doc notes}
%https://docs.google.com/document/d/1Jsi_trBT5cjEXVnGjGHJpcIhMEcVpAn6XmYNxfIHXv0/edit#heading=h.gk1dzxu9wi6q

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
citation ecosystems sub-breakout: \url{http://tinyurl.com/l4tc6a4}
%https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit

@@

\href{http://tinyurl.com/kqpe87z}{Google doc notes}
%https://docs.google.com/document/d/1Mivadj16_9tPrw8Nxmhp72AtNC0BEVO-f7M84FXTbSU/edit#heading=h.elaktnt2tep9

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code Testing \& Code Review}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Karen to lead writing of this section}

\href{http://tinyurl.com/l5t5h45}{Google doc notes}
%https://docs.google.com/document/d/1G1_U5-hkAvDKtTgT4DvUhIYvyYUY4cBBvQHJIdgADa0/edit#heading=h.v7g0nqpn1fur

@@

{\small
\begin{longtable}{ll}
   Jordan Adams          &  Tulane University
\\ Alice Allen           &  Astrophysics Source Code Library (ASCL)
\\ Gabrielle Allen       &  U Illinois
\\ Pierre-Yves Aquilanti &  TOTAL E\&P R\&T USA
\\ Wolfgang Bangerth
\\ David Bernholdt       &  Oak Ridge National Laboratory
\\ Jakob Blomer
\\ Carl Boettiger        &  UCSC/ropensci
\\ Chris Bogart          &  ISR/CMU
\\ Steven R.~Brandt      &  Louisiana State University
\\ Tom Clune             &  NASA GSFC
\\ John W.~Cobb
\\ Dirk Colbry           &  Michigan State University
\\ Karen Cranston        &  NESCent
\\ Tom Crick             &  Cardiff Metropolitan University, UK
\\ Ethan Davis           &  UCAR Unidata
\\ Robert R Downs        &  CIESIN, Columbia University
\\ Anshu Dubey           &  Lawrence Berkeley National Laboratory
\\ Nicole Gasparini      &  Tulane University, New Orleans
\\ Yolanda Gil           &  Information Sciences Institute, University of Southern California
\\ Kurt Glaesemann       &  Pacific northwest national lab
\\ Sol Greenspan         &  National Science Foundation
\\ Ted Habermann         &  The HDF Group
\\ Marcus D.~Hanwell     &  Kitware
\\ Sarah Harris          &  University of Leeds
\\ David Henty           &  EPCC, The University of Edinburgh
\\ Neil Chue Hong        &  Software Sustainability Institute \& University of Edinburgh
\\ James Howison         &  University of Texas
\\ Maxime Hughes
\\ Eric Hutton           &  University of Colorado
\\ Ray Idaszak           &  RENCI/UNC
\\ Samin Ishtiaq         &  Microsoft Research Cambridge, UK
\\ Matt Jones            &  University of California Santa Barbara
\\ Nick Jones            &  New Zealand eScience Infrastructure, University of Auckland
\\ Daniel S.~Katz        &  U Chicago \& Argonne
\\ Ian Kelley
\\ Hilmar Lapp           &  National Evolutionary Synthesis Center (NESCent) and Duke University
\\ Chris Lenhardt
\\ Richard Littauer      &  MIT/University of Saarland
\\ Frank L\"{o}ffler     &  Louisiana State University
\\ Andre Luckow          &  Rutgers
\\ Berkin Malkoc         &  Istanbul Technical University
\\ Kyle Marcus           &  University at Buffalo
\\ Bryan Marker          &  The University of Texas at Austin
\\ Suresh Marru          &  Indiana University
\\ Robert H.~McDonald    &  Indiana University - Data to Insight Center/Libraries
\\ Rupert Nash
\\ Andy Nutter-Upham     &  Whitehead Institute
\\ Abani Patra           &  University at Buffalo
\\ Aleksandra Pawlik     &  Software Sustainability Institute
\\ Cody J.~Permann       &  Idaho National Laboratory
\\ John W.~Peterson      &  Idaho National Laboratory
\\ Benjamin Pharr        &  University of Mississippi
\\ Stephen Piccolo       &  Brigham Young University, Utah
\\ Marlon Pierce         &  Indiana University
\\ Ray Plante            &  NCSA/University of Illinois Urbana-Champaign
\\ Sushil Prasad         &  Georgia State University, Atlanta
\\ Karthik Ram           &  University of California, Berkeley. Berkeley Institute for Data Science and rOpenSci
\\ Mike Rilee            &  NASA/GSFC, Rilee Systems Technologies
\\ Erin Robinson         &  Foundation for Earth Science
\\ Mark Schildhauer      &  NCEAS, Univ. California, Santa Barbara
\\ Jory Schossau         &  Michigan State University
\\ Frank Seinstra        &  Netherlands eScience Center
\\ James Shepherd        &  Rice University
\\ Justin Shi
\\ Ardita Shkurti        &  University of Nottingham
\\ Alan Simpson          &  EPCC, The University of Edinburgh
\\ Carol Song            &  Purdue University
\\ James Spencer         &  Imperial College London
\\ Tracy Teal            &  Data Carpentry
\\ Kaitlin Thaney        &  Mozilla Science Lab
\\ Matt Turk             &  NCSA/UIUC
\\ Colin C.~Venters      &  University of Huddersfield
\\ Nathan Weeks
\\ Ethan White           &  University of Florida/Utah State University
\\ Nancy Wilkins-Diehr   &  SDSC
\\ Greg Wilson           &  Software Carpentry
\end{longtable}
}

>>>>

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\note}[1]{ {\textcolor{blueish}    { ***Note:      #1 }}}
\newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
\newcommand{\gabnote}[1]{ {\textcolor{cyan}    { ***Gabrielle:     #1 }}}
\newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
\newcommand{\manishnote}[1]{  {\textcolor{violet}     { ***Manish: #1 }}}
\newcommand{\davidnote}[1]{  {\textcolor{darkgreen}      { ***David: #1 }}}
\newcommand{\colinnote}[1]{ {\textcolor{red}    {***Colin #1 }}}

@@

\begin{abstract} \todo{TBD due date above}       
\todo{update abstract}
This technical report discusses the submission and peer-review process used by
the Second Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE2) and the results of that process. It is intended to record
both the alternative submission and program organization model used by WSSSPE2
as well as the papers associated with the workshop that resulted from that
process.
\end{abstract}

@@

\note{google doc of notes for reference: \url{http://tinyurl.com/q6ew45v}}
%https://docs.google.com/document/d/1-BxkYWDQ6nNNBXBStUL0xcKF9qCTlEALwf928J_MemI/edit?usp=sharing

@@

This report extends a previous report that discussed the submission, peer-review, and peer-grouping processes in detail.~\cite{WSSSPE2-pre-report}.  It is also based on a collaborative set
of notes taken through a google doc during the workshop.~\cite{WSSSPE2-google-notes} 
Overall, it discusses the organization work done before the workshop (\S\ref{sec:preworkshop}), the keynotes (\S\ref{sec:keynotes}), \todo{discuss rest of report here...} ... exploring sustainability (\S\ref{sec:exploring}) ... code testing and code review (\S\ref{sec:code_testing})

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Submissions, Peer-Review, and Peer-Grouping} \label{sec:preworkshop}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{this section is taken from \cite{WSSSPE2-google-notes}. It could be shortened.}

WSSSPE2 began with a call for papers.
Based on the goal of encouraging a wide range of submissions from those involved
in software practice, ranging from initial thoughts and partial studies to
mature deployments, but focusing on papers that are intended to lead to changes,
the organizers wanted to make submission as easy as possible. The call for
papers stated:

The call included the following areas of interest:
\begin{quote}
\begin{itemize} 
\renewcommand{\labelenumi}{\textbf{\theenumi}.}
\setlength{\rightmargin}{1em}
\item defining software sustainability in the context of science and engineering
software
\begin{itemize}
\item how to evaluate software sustainability
\end{itemize}

@@

@@

Finally, since some of the papers were not represented by cards in the process,
they were not placed in groups by the peer-grouping system, so the authors of
these papers were asked which group seemed the best for their papers, and those
papers were then placed in those groups. Sections~\ref{sec:exploring}-\ref{sec:code_testing}
discuss the breakout groups, including a list of the papers associated with each
group.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes} \label{sec:keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Lead: Choi. Volunteers welcome---contact lead}

@@

The workshop featured two keynote addresses. In the opening keynote
presentation, Kaitlin Thaney of Mozilla Science talked about her organization's
work and policy to enable and support sustainable and reproducible scientific
research. Our second keynote speaker was Neil Chue Hong of Software
Sustainability Institute. He shined the light on how scientific software is
prevalently driving advances in many sciences and engineering fields. Following
the keynotes, participants broke into groups to define and discuss the crucial
notion ``sustainability'' spawned by the speeches of our keynote speakers as well
as the theme of our workshop.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exploring Sustainability} \label{sec:exploring}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Colin to lead writing of this section.}

\note{\href{http://tinyurl.com/mpbhvyb}{Google doc notes}}
%https://docs.google.com/a/uchicago.edu/document/d/10XkshP3YjXnA5JQgP-UFHbOxshWczzehWGSo6V7bNH4/edit

@@

\note{\href{http://tinyurl.com/pn4eq8z}{Google doc notes}}
%https://docs.google.com/a/uchicago.edu/document/d/1_q0jmiEPNFRGAEtGLk_QxQ1vVokgTDYruTGURowNGkY/edit

@@

@@

\note{\href{http://tinyurl.com/k8ruyn9}{Google doc notes}}
%https://docs.google.com/document/d/1Jsi_trBT5cjEXVnGjGHJpcIhMEcVpAn6XmYNxfIHXv0/edit#heading=h.gk1dzxu9wi6q

This group, with just three papers but good interest and participation from attendees,
focused on the institutional, social, and cultural mechanisms encourage the creation
and maintenance of shared software, in the context of what now exists but focusing
on what mechanisms are desired and how we might achieve them.

@@

The group had some initial discussion ...

The group then broke into three subgroups
\begin{itemize}
\item Citation ecosystems
\item Taxonomy of contributorship/Guidelines for software for tenure review
\item Prizes
\end{itemize}

citation ecosystems sub-breakout: \url{http://tinyurl.com/l4tc6a4}
%https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit

@@

@@

\note{\href{http://tinyurl.com/kqpe87z}{Google doc notes}}
%https://docs.google.com/document/d/1Mivadj16_9tPrw8Nxmhp72AtNC0BEVO-f7M84FXTbSU/edit#heading=h.elaktnt2tep9

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code Testing \& Code Review} \label{sec:code_testing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Karen to lead writing of this section}

\note{\href{http://tinyurl.com/l5t5h45}{Google doc notes}}
%https://docs.google.com/document/d/1G1_U5-hkAvDKtTgT4DvUhIYvyYUY4cBBvQHJIdgADa0/edit#heading=h.v7g0nqpn1fur

@@

@@

{\small
\begin{longtable}{ll}
   Jordan Adams          &  Tulane University
\\ Alice Allen           &  Astrophysics Source Code Library (ASCL)
\\ Gabrielle Allen       &  U Illinois
\\ Pierre-Yves Aquilanti &  TOTAL E\&P R\&T USA
\\ Wolfgang Bangerth & Texas A\&M
\\ David Bernholdt       &  Oak Ridge National Laboratory
\\ Jakob Blomer
\\ Carl Boettiger        &  UCSC/ropensci
\\ Chris Bogart          &  ISR/CMU
\\ Steven R.~Brandt      &  Louisiana State University
\\ Tom Clune             &  NASA GSFC
\\ John W.~Cobb
\\ Dirk Colbry           &  Michigan State University
\\ Karen Cranston        &  NESCent
\\ Tom Crick             &  Cardiff Metropolitan University, UK
\\ Ethan Davis           &  UCAR Unidata
\\ Robert R Downs        &  CIESIN, Columbia University
\\ Anshu Dubey           &  Lawrence Berkeley National Laboratory
\\ Nicole Gasparini      &  Tulane University, New Orleans
\\ Yolanda Gil           &  Information Sciences Institute, University of Southern California
\\ Kurt Glaesemann       &  Pacific northwest national lab
\\ Sol Greenspan         &  National Science Foundation
\\ Ted Habermann         &  The HDF Group
\\ Marcus D.~Hanwell     &  Kitware
\\ Sarah Harris          &  University of Leeds
\\ David Henty           &  EPCC, The University of Edinburgh
\\ Neil Chue Hong        &  Software Sustainability Institute \& University of Edinburgh
\\ James Howison         &  University of Texas
\\ Maxime Hughes
\\ Eric Hutton           &  University of Colorado
\\ Ray Idaszak           &  RENCI/UNC
\\ Samin Ishtiaq         &  Microsoft Research Cambridge, UK
\\ Matt Jones            &  University of California Santa Barbara
\\ Nick Jones            &  New Zealand eScience Infrastructure, University of Auckland
\\ Daniel S.~Katz        &  U Chicago \& Argonne
\\ Ian Kelley
\\ Hilmar Lapp           &  National Evolutionary Synthesis Center (NESCent) and Duke University
\\ Chris Lenhardt
\\ Richard Littauer      &  MIT/University of Saarland
\\ Frank L\"{o}ffler     &  Louisiana State University
\\ Andre Luckow          &  Rutgers
\\ Berkin Malkoc         &  Istanbul Technical University
\\ Kyle Marcus           &  University at Buffalo
\\ Bryan Marker          &  The University of Texas at Austin
\\ Suresh Marru          &  Indiana University
\\ Robert H.~McDonald    &  Indiana University - Data to Insight Center/Libraries
\\ Rupert Nash
\\ Andy Nutter-Upham     &  Whitehead Institute
\\ Abani Patra           &  University at Buffalo
\\ Aleksandra Pawlik     &  Software Sustainability Institute
\\ Cody J.~Permann       &  Idaho National Laboratory
\\ John W.~Peterson      &  Idaho National Laboratory
\\ Benjamin Pharr        &  University of Mississippi
\\ Stephen Piccolo       &  Brigham Young University, Utah
\\ Marlon Pierce         &  Indiana University
\\ Ray Plante            &  NCSA/University of Illinois Urbana-Champaign
\\ Sushil Prasad         &  Georgia State University, Atlanta
\\ Karthik Ram           &  University of California, Berkeley. Berkeley Institute for Data Science and rOpenSci
\\ Mike Rilee            &  NASA/GSFC, Rilee Systems Technologies
\\ Erin Robinson         &  Foundation for Earth Science
\\ Mark Schildhauer      &  NCEAS, Univ. California, Santa Barbara
\\ Jory Schossau         &  Michigan State University
\\ Frank Seinstra        &  Netherlands eScience Center
\\ James Shepherd        &  Rice University
\\ Justin Shi
\\ Ardita Shkurti        &  University of Nottingham
\\ Alan Simpson          &  EPCC, The University of Edinburgh
\\ Carol Song            &  Purdue University
\\ James Spencer         &  Imperial College London
\\ Tracy Teal            &  Data Carpentry
\\ Kaitlin Thaney        &  Mozilla Science Lab
\\ Matt Turk             &  NCSA/UIUC
\\ Colin C.~Venters      &  University of Huddersfield
\\ Nathan Weeks
\\ Ethan White           &  University of Florida/Utah State University
\\ Nancy Wilkins-Diehr   &  SDSC
\\ Greg Wilson           &  Software Carpentry
\end{longtable}
}
